---
title: "Rectal Upfront Surgery"
author: "Dave Hein"
date: "2/7/2020"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=8, fig.height=4)
library(tidyverse)
library(haven)
library(lmtest)
library(glmnet)
library(nnet)
```


### Selecting variables from NCDB data base that are useful
```{r}
#Reading in data

Rectal_Data <- read_dta("C:/Users/Dave Work/OneDrive - University of Texas Southwestern/UpFrontSurgery/020320 Surg Alone Rectal.dta")
 
Rectal_Data <- Rectal_Data %>% select(
                                           RX_SUMM_SYSTEMIC_SUR_SEQ,RX_SUMM_RADIATION, RX_SUMM_SURGRAD_SEQ,
                                           RX_SUMM_CHEMO,  SEX,MED_INC_QUAR_12, AGE,
                                           PUF_CASE_ID, raceb, TNM_CLIN_STAGE_GROUP,TNM_PATH_STAGE_GROUP,SEQUENCE_NUMBER, HISTOLOGY, 
                                           RX_SUMM_SURG_PRIM_SITE, YEAR_OF_DIAGNOSIS, FACILITY_TYPE_CD, INSURANCE_STATUS, CDCC_TOTAL_BEST, 
                                           DX_DEFSURG_STARTED_DAYS, CLASS_OF_CASE
)

#getting rid of NAs and Nulls
df1 <- na.omit(Rectal_Data )
```


### Cleaning up data
```{r}
# Drop if information regarding radiation, chemo, or surgery is unknown, missing, or if radiation was ineroperative
df1 <- df1[(
              df1$RX_SUMM_SYSTEMIC_SUR_SEQ != 6 & df1$RX_SUMM_SYSTEMIC_SUR_SEQ != 7 & df1$RX_SUMM_SYSTEMIC_SUR_SEQ != 9 &
              df1$RX_SUMM_RADIATION != 9 & df1$RX_SUMM_SURGRAD_SEQ != 6 & df1$RX_SUMM_SURGRAD_SEQ != 9 & df1$RX_SUMM_CHEMO != 88 &
              df1$RX_SUMM_CHEMO != 99
          ),]

#drop if not stage 2 or 3 
df1 <- df1 %>% mutate(stage3 = ifelse(str_detect(TNM_CLIN_STAGE_GROUP,"2"),"2","0")) 
df1$stage3[str_detect(df1$TNM_CLIN_STAGE_GROUP,"3")] <- "3"
df1 <- df1[!(df1$stage3 =="0"),]

#Drop if not first primary 
df1 <- df1%>%filter(SEQUENCE_NUMBER %in% c("00","01"))

#drop if histology not adenocarcinoma 
df1 <- df1%>%filter(HISTOLOGY %in% c(8140,8210,8260,8261,8262,8263,8480,8481))

#drop if did not get defin surg 
df1 <- df1%>%filter(RX_SUMM_SURG_PRIM_SITE %in% c(40,30,50,60,70,80))

```


### Assigning new variables 
```{r}
#Assigning neo adjuvant if got rad OR chemo before surg
df1$surg4 <- NA
df1$surg4[df1$RX_SUMM_SURGRAD_SEQ == 2 | df1$RX_SUMM_SURGRAD_SEQ == 4] <- 0
df1$surg4[df1$RX_SUMM_SYSTEMIC_SUR_SEQ == 2 | df1$RX_SUMM_SYSTEMIC_SUR_SEQ == 4] <- 0

df1$surg4[df1$RX_SUMM_SYSTEMIC_SUR_SEQ==3 & df1$RX_SUMM_RADIATION==0] <- 1
df1$surg4[df1$RX_SUMM_SURGRAD_SEQ==3 & df1$RX_SUMM_CHEMO==0] <- 1
df1$surg4[df1$RX_SUMM_SURGRAD_SEQ==3 & df1$RX_SUMM_CHEMO==82] <- 1
df1$surg4[df1$RX_SUMM_SURGRAD_SEQ==3 & df1$RX_SUMM_CHEMO==85] <- 1
df1$surg4[df1$RX_SUMM_SURGRAD_SEQ==3 & df1$RX_SUMM_CHEMO==86] <- 1
df1$surg4[df1$RX_SUMM_SURGRAD_SEQ==3 & df1$RX_SUMM_CHEMO==87] <- 1
df1$surg4[df1$RX_SUMM_SURGRAD_SEQ==3 & df1$RX_SUMM_SYSTEMIC_SUR_SEQ==3] <- 1
df1$surg4[df1$RX_SUMM_RADIATION==0 & df1$RX_SUMM_CHEMO==0] <- 1
df1$surg4[df1$RX_SUMM_RADIATION==0 & df1$RX_SUMM_CHEMO==82] <- 1
df1$surg4[df1$RX_SUMM_RADIATION==0 & df1$RX_SUMM_CHEMO==85] <- 1
df1$surg4[df1$RX_SUMM_RADIATION==0 & df1$RX_SUMM_CHEMO==86] <- 1
df1$surg4[df1$RX_SUMM_RADIATION==0 & df1$RX_SUMM_CHEMO==87] <- 1

df1 <- na.omit(df1)

#age group under 45, 45-65, 65 and up AGE
df1$agegroup <- NA
df1$agegroup[df1$AGE<=45]<-0
df1$agegroup[df1$AGE>=46]<-1
df1$agegroup[df1$AGE>=65]<-2

#people who switch facility
df1 <- df1 %>% mutate(treatnotatdiag = ifelse(CLASS_OF_CASE %in% c(0,22,20,21),"Switched","Stayed"))

```


### Logistic regression for odds of upfront surgery
```{r}
#making variables factors
df1[] <- lapply(df1, factor)
df1$stage3 <- relevel(df1$stage3, ref = "3")
   
#regression
fitlog <- glm(surg4 ~ YEAR_OF_DIAGNOSIS + MED_INC_QUAR_12  +INSURANCE_STATUS + raceb + CDCC_TOTAL_BEST + SEX + agegroup + stage3 + FACILITY_TYPE_CD + treatnotatdiag,  data = df1, family = binomial(link = logit))

#significance
summary(fitlog)
#AOR and 95% CI
    #exp(cbind(OR = coef(fitlog), confint(fitlog)))
```

		
### Graphing proportion receiving upfront surg at different facility types over time
```{r}
#Make surg4 numeric again so we can calculate a ratio in code below
df1<-df1%>%mutate(surg4=as.numeric(as.character(surg4)))

#renaming the fac type
dfyear <- df1 %>% dplyr::select(YEAR_OF_DIAGNOSIS, surg4, FACILITY_TYPE_CD)
dfyear$Facility_Type<-NA
dfyear$Facility_Type[dfyear$FACILITY_TYPE_CD == 1] <- "Community Cancer Program"
dfyear$Facility_Type[dfyear$FACILITY_TYPE_CD == 2] <- "Comprehensive CC Program"
dfyear$Facility_Type[dfyear$FACILITY_TYPE_CD == 3] <- "Academic Program"
dfyear$Facility_Type[dfyear$FACILITY_TYPE_CD == 4] <- "Integrated Network Cancer Program"

#finding proportion of upfront surg by fac type

dfyear <- dfyear %>% dplyr::group_by(YEAR_OF_DIAGNOSIS,Facility_Type) %>% dplyr::summarize(ratio=mean(surg4))

#finding overall median per year of upfront surg
dfyearavg <- df1 %>% dplyr::group_by(YEAR_OF_DIAGNOSIS) %>% dplyr::summarize(ratio = mean(surg4))
dfyearavg$Facility_Type <- "Overall Median"
dfyearavg <- dfyearavg %>% select(YEAR_OF_DIAGNOSIS,Facility_Type,ratio)

#Merging overall and by facility type
year1 <- as.data.frame(dfyear)
year2 <- as.data.frame(dfyearavg)
mergedyearly <- rbind(year1,year2)

#plotting upfron surg by fac type
ggplot() + geom_line(data = mergedyearly, aes(x=YEAR_OF_DIAGNOSIS, y=ratio, group = Facility_Type,color=Facility_Type),size = 1.8) + labs(x="Year of Diagnosis", y = "Proportion Receiving Upfront Surgery", color = "Facility Type") +theme_test() +scale_color_brewer(palette = "Dark2") 
```


### Looking at how delays in surgery and facility switching are linked (note some data points are cut off in graph)
```{r}

#Making df of just people who got upfront surg
df2 <- df1 %>% filter(surg4 ==1)

#making three groups for a surgery delay
df2<-df2%>%mutate(DX_DEFSURG_STARTED_DAYS=as.numeric(DX_DEFSURG_STARTED_DAYS))
df2$surggroup<-NA
df2$surggroup[df2$DX_DEFSURG_STARTED_DAYS == 0] <- "Day zero"
df2$surggroup[df2$DX_DEFSURG_STARTED_DAYS > 0 & df2$DX_DEFSURG_STARTED_DAYS < 28] <- "Short delay"
df2$surggroup[df2$DX_DEFSURG_STARTED_DAYS >= 28] <- "Long Delay"
df2$surggroup <- factor(df2$surggroup)  
  
#plotting surg delay and switching facility
ggplot(df2,aes(x=DX_DEFSURG_STARTED_DAYS,fill=treatnotatdiag))+geom_histogram(binwidth = 5) + xlim(-10,200) + labs(x="Days Untill Surgery From Diagnosis", y = "Count", fill = "Facility Switch")+theme_test() + scale_fill_brewer(palette = "Set2") + geom_vline(xintercept=28) 
```
  
  
### Multinomial logistic regression 
```{r}
#logistic regression for facoters involved in surg delay
test <- multinom(surggroup ~  CDCC_TOTAL_BEST  + FACILITY_TYPE_CD + treatnotatdiag, data = df2)
summary(test)
#z <- summary(test)$coefficients/summary(test)$standard.errors
#p <- (1 - pnorm(abs(z), 0, 1)) * 2
#p
#exp(coef(test))
#exp(confint(test))
```




